---
title: "Analysis of Cellular Proteomic Processes"
date: "December 8, 2022"
subtitle: "Keerat Gill, Michelle Wang, Kefira Wang, Ali Hafeez"
output: 
  beamer_presentation:
    theme: Hannover
    colortheme: beaver
    fonttheme: structureitalicserif
    slide_level: 2
classoption: aspectratio=169
fontsize: 9pt
urlcolor: blue
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
```


## Introduction

- The data is based on Flow Cytometry for single-cell analysis and Mass Spectrometry for measurement of cellular proteomic processes.\
- Measures 22 AP-1 transcription factors and 4 phenotype proteins.\
What is *"good"* cellular homeostasis? **And can bad proteins become good?**\

\vspace{0.5cm}

Our main research questions are:\
1. Can we predict cellular phenotypical outcomes from transcription factors?\
2. Do the phenotype levels change over time in different drug conditions?\
3. What is the relationship between different proteins at a fixed time?\

\vspace{0.5cm}

### Statistical Methods
- We first omitted all NA values to make sure our data was complete.\
- We designated each phenotype a HIGH or LOW classification depending on if the recorded value was higher than the mean.\
- Using the HIGH/LOW designation, we designated each observation to be melanocytic or undifferentiated.\

## AXL Classification Tree 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
big_data <- read_csv("STA130Project.csv") #uploading the data
clean_data <- na.omit(big_data)

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
clean_data_class <- clean_data %>% mutate(NGFR_class = ifelse(NGFR >= mean(NGFR), 'HIGH', 'LOW'), AXL_class = ifelse(AXL >= mean(AXL), 'HIGH', 'LOW'), Sox10_class = ifelse(Sox10 >= mean(Sox10), 'HIGH', 'LOW'), MiTFg_class = ifelse(MiTFg >= mean(MiTFg), 'HIGH', 'LOW'))
#glimpse(clean_data_class)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
clean_data_class_only_TFs <- clean_data_class %>% select(-timepoint_id, -Timepoint, -drug_id, -Drugs, -dose_id, -Doses, -Rep)
#glimpse(clean_data_class_only_TFs)
TFs <- clean_data_class_only_TFs %>% select(-MiTFg, -MiTFg_class, -AXL, -AXL_class, -NGFR, -NGFR_class, -Sox10, -Sox10_class)
# TFs_names: Phospho_c_Fos + Phospho_c_Jun + Phospho_ATF2 + Phospho_Fra1 + c_Fos + c_Jun + Fra1 + JunD + ATF2 + JunB + Fra2 + ATF4 + Phospho_ATF4 + Phospho_Erk1 + Phospho_ATF1 + ATF6 + Phospho_S6 + ATF3 + ATF5 + ATF6 + Phospho_S6 + ATF3 + ATF5 + Phospho_p38 + Ki_67 + NF_kappaB
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.align='center'}
library(rpart)
library(partykit)
AXL_tree <- rpart(AXL_class ~ Phospho_c_Fos + Phospho_c_Jun + Phospho_ATF2 + Phospho_Fra1 + c_Fos + c_Jun + Fra1 + JunD + ATF2 + JunB + Fra2 + ATF4 + Phospho_ATF4 + Phospho_Erk1 + Phospho_ATF1 + ATF6 + Phospho_S6 + ATF3 + ATF5 + ATF6 + Phospho_S6 + ATF3 + ATF5 + Phospho_p38 + Ki_67 + NF_kappaB, data = clean_data_class_only_TFs)
plot(as.party(AXL_tree), type='simple', gp=gpar(cex = 0.6), main = 'Classification Tree to Determine AXL Class')
```

## NGFR Classification Tree

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.align='center'}
NGFR_tree <- rpart(NGFR_class ~ Phospho_c_Fos + Phospho_c_Jun + Phospho_ATF2 + Phospho_Fra1 + c_Fos + c_Jun + Fra1 + JunD + ATF2 + JunB + Fra2 + ATF4 + Phospho_ATF4 + Phospho_Erk1 + Phospho_ATF1 + ATF6 + Phospho_S6 + ATF3 + ATF5 + ATF6 + Phospho_S6 + ATF3 + ATF5 + Phospho_p38 + Ki_67 + NF_kappaB, data = clean_data_class_only_TFs)
plot(as.party(NGFR_tree), type='simple', main='Classification Tree to Determine NGFR Class', gp=gpar(cex=0.4))
```

## MiTFg Classification Tree

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.align='center'}
MiTFg_tree <- rpart(MiTFg_class ~ Phospho_c_Fos + Phospho_c_Jun + Phospho_ATF2 + Phospho_Fra1 + c_Fos + c_Jun + Fra1 + JunD + ATF2 + JunB + Fra2 + ATF4 + Phospho_ATF4 + Phospho_Erk1 + Phospho_ATF1 + ATF6 + Phospho_S6 + ATF3 + ATF5 + ATF6 + Phospho_S6 + ATF3 + ATF5 + Phospho_p38 + Ki_67 + NF_kappaB, data = clean_data_class_only_TFs)
plot(as.party(MiTFg_tree), type='simple', main = 'Classification Tree to Determine MiTFg Class', gp=gpar(cex=0.6))
```

## Sox10 Classification Tree

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.align='center'}
Sox10_tree <- rpart(Sox10_class ~ Phospho_c_Fos + Phospho_c_Jun + Phospho_ATF2 + Phospho_Fra1 + c_Fos + c_Jun + Fra1 + JunD + ATF2 + JunB + Fra2 + ATF4 + Phospho_ATF4 + Phospho_Erk1 + Phospho_ATF1 + ATF6 + Phospho_S6 + ATF3 + ATF5 + ATF6 + Phospho_S6 + ATF3 + ATF5 + Phospho_p38 + Ki_67 + NF_kappaB, data = clean_data_class_only_TFs)
plot(as.party(Sox10_tree), type='simple', main='Classification Tree to Determine Sox10 Class', gp=gpar(cex=0.6))
```

## Classification Tree Findings

### Findings from Classification Trees - Cancerous Cells (Melanocytic)
AXL: LOW\
- Phospho_Fra1 < 2.834\
- Phospho_Fra1 >= 2.834, Fra2 < 2.722 AND Phospho_Erk1 >= 3.361\
- Phospho_Fra1 >= 2.834, Fra2 < 2.722, Phospho_Erk1 < 3.361 AND Fra2 < 2.533\
NGFR: HIGH\
- Phospho_ATF2 < 2.362 AND ATF6 >= 2.749 AND Ki_67 < 3.739\
- Phospho_ATF2 >= 2.362 AND ATF6 >= 2.795\
- Phospho_ATF2 >= 2.362 AND ATF6 < 2.795 AND Ki_67 < 3.68 AND c_Fos >= 2.417\
MiTFg: HIGH\
- ATF6 >= 2.736 AND ATF6 >= 2.833\
- ATF6 >= 2.736 AND ATF6 < 2.833 AND Fra2 < 2.747\
- ATF6 >= 2.736 AND ATF6 < 2.833 AND Fra2 >= 2.747 AND ATF4 >= 2.795\
SOX10: LOW\
- Phospho_S6 >= 3.331 AND ATF4 < 2.767\
- Phospho_S6 >= 3.331, ATF4 >= 2.767, ATF3 < 3.53 AND Phospho_Erk1 >= 3.337\

---

### Findings from Classification Trees - Non-Cancerous Cells (Undifferentiated)

AXL: HIGH\
- Phospho_Fra1 >= 2.834 AND Fra2 >= 2.722\
- Phospho_Fra1 >= 2.834, Fra2 < 2.722, Phospho_Erk1 < 3.361 AND Fra2 >= 2.533\
NGFR: LOW\
- Phospho_ATF2 < 2.362 AND ATF6 < 2.749\
- Phospho_ATF2 < 2.362 AND ATF6 >= 2.749 AND Ki_67 >= 3.739\
- Phospho_ATF2 >= 2.362 AND ATF6 < 2.795 AND Ki_67 >= 3.68\
- Phospho_ATF2 >= 2.362 AND ATF6 < 2.795 AND Ki_67 < 3.68 AND c_Fos < 2.417\
MiTFg: LOW\
- ATF6 < 2.736\
- ATF6 >= 2.736 AND ATF6 < 2.833 AND Fra2 >= 2.747 AND ATF4 < 2.795\
SOX10: LOW\
- Phospho_S6 >= 3.331 AND ATF4 < 2.767\
- Phospho_S6 >= 3.331, ATF4 >= 2.767, ATF3 < 3.53, Phospho_Erk1 >= 3.337\


## Turning Bad Cells ino Good Cells

**Do the phenotype levels change over time in different drug conditions?**\

\vspace{0.5cm}

- We first plotted the average level of phenotypes in melanocytic cells over time, grouped by drug type, to see if there was a noticable difference to justify a two-sample hypothesis test.\
- NGFR, AXL and MiTFg had noticable  linear trends in their means overtime.\
- Sox10 had an abnormal jump at the 2h timepoint for both drugs.\
- This was taken into consideration when conducting the two-sample test.\
- The following are the average means and p-values for the change in phenotype level between the 0.5h and 24h timepoint.\
- Alpha-significance level is 0.05, as per scientific convention.

\vspace{0.5cm}

*NULL HYPOTHESIS: There is no change in protein level over time.*

---

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#creating new variables NGFR_class, AXL_class, Sox10_class and MiTFg_class to later use to create cancerous and non_cancerous subsets.

cleandata_class <- clean_data %>% mutate(NGFR_class = ifelse(NGFR >= mean(NGFR), 'HIGH', 'LOW'), 
                     AXL_class = ifelse(AXL >= mean(AXL), 'HIGH', 'LOW'), 
                     Sox10_class = ifelse(Sox10 >= mean(Sox10),'HIGH', 'LOW'),
                     MiTFg_class = ifelse(MiTFg >= mean(MiTFg), 'HIGH', 'LOW'))

cancerous <- cleandata_class %>% filter(AXL_class == 'LOW' & Sox10_class == 'LOW' & NGFR_class == 'HIGH' & MiTFg_class == 'HIGH')

```

```{r, echo=FALSE, message=FALSE, warning=FALSE,}

#turning Timepoint into a numerical data by creating new variable, orderedtime.
#SPECIFICALLY ONLY FOR CANCEROUS CELLS

groupcancer_time <- cancerous %>% mutate(orderedtime = case_when(
                                                         Timepoint == '0.5 h' ~ 0.5, 
                                                         Timepoint == '2 h' ~ 2,
                                                         Timepoint == '6 h' ~ 6,
                                                         Timepoint == '15 h' ~ 15,
                                                         Timepoint == '24 h' ~ 24))

groupcancer <- groupcancer_time %>% select(Drugs, NGFR, AXL, Sox10, MiTFg, NGFR_class, AXL_class, Sox10_class, MiTFg_class, orderedtime)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE,}

##Creating sub-data sets within the cancerous cells that separate cells based on which drug they are using and their timepoint (time of recorded death).

#Notation; C - cancerous, V - Vem drug, VT - Vem+Tram drug. '___time' - indicates the time. 

#This one has only VEM drugs 

CVpointfivetime <- groupcancer %>% filter(Drugs == 'Vem' & orderedtime == 0.5)

CVtwotime <- groupcancer %>% filter(Drugs == 'Vem' & orderedtime == 2)

CVsixtime <- groupcancer %>% filter(Drugs == 'Vem' & orderedtime == 6)

CVfifteentime <- groupcancer %>% filter(Drugs == 'Vem' & orderedtime == 15)

CVfulltime <- groupcancer %>% filter(Drugs == 'Vem' & orderedtime == 24)

##This one only has Vem+Tram drugs

CVTpointfivetime <- groupcancer %>% filter(Drugs == 'Vem+Tram' & orderedtime == 0.5)

CVTtwotime <- groupcancer %>% filter(Drugs == 'Vem+Tram' & orderedtime == 2)

CVTsixtime <- groupcancer %>% filter(Drugs == 'Vem+Tram' & orderedtime == 6)

CVTfifteentime <- groupcancer %>% filter(Drugs == 'Vem+Tram' & orderedtime == 15)

CVTfulltime <- groupcancer %>% filter(Drugs == 'Vem+Tram' & orderedtime == 24)

```

## AXL Levels Change Over Time in VEM and VEM+TRAM Drug

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=2.5}

#Creating subset of melanocytic cells, only looking at the average level of AXL at each time point. Specifically for VEM drug.

AXLVemC <- groupcancer %>% filter(Drugs == 'Vem') %>% select(AXL, orderedtime) %>% group_by(orderedtime) %>% summarize(n=n(), mean(AXL)) 

AXLtimeCV <- c(0.5, 2, 6, 15, 24)
AXLmeanCV <- c(3.366909, 3.333558, 3.335184, 3.318167, 3.301639) #Taken from AXLVemC

AXLmean_df_CV <- data.frame(AXLtimeCV, AXLmeanCV)

ggplot(AXLmean_df_CV, aes(x=AXLtimeCV, y=AXLmeanCV)) + geom_point() + labs(title='AXL in VEM drug', x='time', y= 'AXL level')

#cor(AXLmean_df_CV$AXLtimeCV, AXLmean_df_CV$AXLmeanCV)

#Do the same thing, with the VEM+TRAM drug

AXLVTramC <- groupcancer %>% filter(Drugs == 'Vem+Tram') %>% select(AXL, orderedtime) %>% group_by(orderedtime) %>% summarize(n=n(), mean(AXL)) 


AXLtimeCVT <- c(0.5, 2, 6, 15, 24)
AXLmeanCVT <- c(3.357275, 3.337623, 3.327603,	3.308636,	3.316986) #Taken from AXLVTramC

AXLmean_df_CVT <- data.frame(AXLtimeCVT, AXLmeanCVT) 

ggplot(AXLmean_df_CVT, aes(x=AXLtimeCVT, y=AXLmeanCVT)) + geom_point() + labs(title='AXL in V+T drug', x='time', y= 'AXL level')

#cor(AXLmean_df_CVT$AXLtimeCVT, AXLmean_df_CVT$AXLmeanCVT)

```

## Sox10 Level Change Over Time in VEM and VEM+TRAM Drug

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=2.5}

#Creating subset of melanocytic cells, only looking at the average level of Sox10 at each time point. Specifically for VEM drug.

Sox10VemC <- groupcancer %>% filter(Drugs == 'Vem') %>% select(Sox10, orderedtime) %>% group_by(orderedtime) %>% summarize(n=n(), mean(Sox10)) 


Sox10timeCV <- c(0.5, 2, 6, 15, 24)
Sox10meanCV <- c(3.546383, 3.406140, 3.543998, 3.543643, 3.537207)

Sox10mean_df_CV <- data.frame(Sox10timeCV, Sox10meanCV)


ggplot(Sox10mean_df_CV, aes(x=Sox10timeCV, y=Sox10meanCV)) + geom_point() + labs(title='Sox10 in Vem drug', x='time', y= 'Sox10 level')

#cor(Sox10mean_df_CV$Sox10timeCV, Sox10mean_df_CV$Sox10meanCV)

#Same thing, with VEM+TRAM Drug

Sox10VTramC <- groupcancer %>% filter(Drugs == 'Vem+Tram') %>% select(Sox10, orderedtime) %>% group_by(orderedtime) %>% summarize(n=n(), mean(Sox10)) 

Sox10timeCVT <- c(0.5, 2, 6, 15, 24)
Sox10meanCVT <- c(3.548716, 3.435378, 3.544271, 3.547409, 3.539588)	

Sox10mean_df_CVT <- data.frame(Sox10timeCVT, Sox10meanCVT)

ggplot(Sox10mean_df_CVT, aes(x=Sox10timeCVT, y=Sox10meanCVT)) + geom_point() + labs(title= 'Sox10 in V+T drug', x='time', y= 'Sox10 level')

#cor(Sox10mean_df_CVT$Sox10timeCVT, Sox10mean_df_CVT$Sox10meanCVT)

```


## NGFR Level Change Over Time in VEM and VEM+TRAM Drug

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=2.5}

#Creating subset of melanocytic cells, only looking at the average level of NGFR at each time point. Specifically for VEM drug.

NGFRVemC <- groupcancer %>% filter(Drugs == 'Vem') %>% select(NGFR, orderedtime) %>% group_by(orderedtime) %>% summarize(n=n(), mean(NGFR)) 


NGFRtimeCV <- c(0.5, 2, 6, 15, 24)
NGFRmeanCV <- c(3.105181, 3.122855, 3.121872, 3.123155, 3.131846)	#Taken from NGFRVemC
  
NGFRmean_df_CV <- data.frame(NGFRtimeCV, NGFRmeanCV)

ggplot(NGFRmean_df_CV, aes(x=NGFRtimeCV, y=NGFRmeanCV)) + geom_point() + labs(title='NGFR in Vem drug', x='time', y= 'NGFR level')

#cor(NGFRmean_df_CV$NGFRtimeCV, NGFRmean_df_CV$NGFRmeanCV)

#Do the same thing with VEM+TRAM Drug

NGFRVTramC <- groupcancer %>% filter(Drugs == 'Vem+Tram') %>% select(NGFR, orderedtime) %>% group_by(orderedtime) %>% summarize(n=n(), mean(NGFR)) 

NGFRtimeCVT <- c(0.5, 2, 6, 15, 24)
NGFRmeanCVT <- c(3.111332, 3.122884,	3.117180, 3.125491, 3.145561) #Taken from NGFRVTramC
  
NGFRmean_df_CVT <- data.frame(NGFRtimeCVT, NGFRmeanCVT)


ggplot(NGFRmean_df_CVT, aes(x=NGFRtimeCVT, y=NGFRmeanCVT)) + geom_point() + labs(title='NGFR in V+T drug', x='time', y= 'NGFR level')

#cor(NGFRmean_df_CVT$NGFRtimeCVT, NGFRmean_df_CVT$NGFRmeanCVT)


```

## MiTFg Level Change Over Time in VEM and VEM+TRAM Drug


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=2.5}

#Creating subset of melanocytic cells, only looking at the average level of MiTFg at each time point. Specifically for VEM drug.

MiTFgVemC <- groupcancer %>% filter(Drugs == 'Vem') %>% select(MiTFg, orderedtime) %>% group_by(orderedtime) %>% summarize(n=n(), mean(MiTFg)) 

MiTFgtimeCV <- c(0.5, 2, 6, 15, 24)
MiTFgmeanCV <- c(3.105181, 3.122855, 3.121872, 3.123155, 3.131846)	#Taken from MiTFgVemC
  
MiTFgmean_df_CV <- data.frame(MiTFgtimeCV, MiTFgmeanCV)

ggplot(MiTFgmean_df_CV, aes(x=MiTFgtimeCV, y=MiTFgmeanCV)) + geom_point() + labs(title='MiTFg in Vem drug', x='time', y= 'MiTFg level')

#cor(MiTFgmean_df_CV$MiTFgtimeCV, MiTFgmean_df_CV$MiTFgmeanCV)

#Do the same for VEM+TRAM DRUG

MiTFgVTramC <- groupcancer %>% filter(Drugs == 'Vem+Tram') %>% select(MiTFg, orderedtime) %>% group_by(orderedtime) %>% summarize(n=n(), mean(MiTFg)) 

MiTFgtimeCVT <- c(0.5, 2, 6, 15, 24)
MiTFgmeanCVT <- c(3.835648,	3.832574, 3.803088,	3.784568,	3.782575) #TAken from MiTFgVTramC
  
MiTFgmean_df_CVT <- data.frame(MiTFgtimeCVT, MiTFgmeanCVT)

ggplot(MiTFgmean_df_CVT, aes(x=MiTFgtimeCVT, y=MiTFgmeanCVT)) + geom_point() + labs(title='MiTFg in V+T drug', x='time', y= 'MiTFg level')

#cor(MiTFgmean_df_CVT$MiTFgtimeCVT, MiTFgmean_df_CVT$MiTFgmeanCVT)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Two sample hypothesis testing AXL, VEM drug

AXLstarttimeCV <- CVpointfivetime %>% select(AXL) #at 0.5 hours, #only 138 data points
AXLfulltimeCV <- CVfulltime %>% select(AXL) #at 14 hours, has 5,000 data points

n1 <- 138
n2 <- 4000

set.seed(523)
N <- 1000; #how many times you run the code
AXLsimulated_values <- 1:N

for(i in 1:N){ #do 1000 'fake' observations
  AXLshuffled_CV <- sample(c(AXLstarttimeCV$AXL, AXLfulltimeCV$AXL), size=n1+n2, replace=FALSE)
  AXLsimvalue_C <- mean(AXLshuffled_CV[1:n1])-mean(AXLshuffled_CV[(n1+1):(n1+n2)])
  AXLsimulated_values[i] <- as.numeric(AXLsimvalue_C)
}

simAXLCV <- tibble(AXLmean_diff = AXLsimulated_values)

#simAXLCV %>% ggplot(aes(x=AXLmean_diff)) + geom_histogram(binwidth=0.001, color="black", fill="gray")

AXLtest_stat <- 3.366909 - 3.301639

AXL_moreextreme <- simAXLCV %>% filter(abs(AXLmean_diff) >= abs(AXLtest_stat)) %>% summarise(n()/100)

AXL_p_CV <- as.numeric(AXL_moreextreme); #AXL_p_CV

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Two sample hypothesis testing AXL, VEM+TRAM drug

AXLstarttimeCVT <- CVTpointfivetime %>% select(AXL) #at 0.5 hours, #only 138 data points
AXLfulltimeCVT <- CVTfulltime %>% select(AXL) #at 14 hours, has 5,000 data points

n1 <- 195
n2 <- 2194

set.seed(523)
N <- 1000; #how many times you run the code
AXLsimulated_valuesCVT <- 1:N

for(i in 1:N){ #do 1000 'fake' observations
  AXLshuffled_CVT <- sample(c(AXLstarttimeCVT$AXL, AXLfulltimeCVT$AXL), size=n1+n2, replace=FALSE)
  AXLsimvalue_CVT <- mean(AXLshuffled_CVT[1:n1])-mean(AXLshuffled_CVT[(n1+1):(n1+n2)])
  AXLsimulated_valuesCVT[i] <- as.numeric(AXLsimvalue_CVT)
}

simAXLCVT <- tibble(AXLmean_diff_CVT = AXLsimulated_valuesCVT)

#simAXLCVT %>% ggplot(aes(x=AXLmean_diff_CVT)) + geom_histogram(binwidth=0.001, color="black", fill="gray")

AXLtest_stat_CVT <- 3.357275 - 3.316986

AXL_moreextreme_CVT <- simAXLCVT %>% filter(abs(AXLmean_diff_CVT) >= abs(AXLtest_stat_CVT)) %>% summarise(n()/100)

AXL_p_CVT <- as.numeric(AXL_moreextreme_CVT); #AXL_p_CVT

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Two sample hypothesis testing SOX, VEM drug

SOXstarttimeCV <- CVpointfivetime %>% select(Sox10) #at 0.5 hours, #only 138 data points
SOXfulltimeCV <- CVfulltime %>% select(Sox10) #at 24 hours, has 5,000 data points

n1 <- 135
n2 <- 4000

set.seed(523)
N <- 1000; #how many times you run the code
SOXsimulated_values_CV <- 1:N

for(i in 1:N){ #do 1000 'fake' observations
  SOXshuffled_CV <- sample(c(SOXstarttimeCV$Sox10, SOXfulltimeCV$Sox10), size=n1+n2, replace=FALSE)
  SOXsimvalue_CV <- mean(SOXshuffled_CV[1:n1])-mean(SOXshuffled_CV[(n1+1):(n1+n2)])
  SOXsimulated_values_CV[i] <- as.numeric(SOXsimvalue_CV)
}

simSOX <- tibble(SOXmean_diff = SOXsimulated_values_CV)

#simSOX %>% ggplot(aes(x=SOXmean_diff)) + geom_histogram(binwidth=0.001, color="black", fill="gray")

SOXtest_stat <- 3.546383 - 3.537207

SOX_moreextreme <- simSOX %>% filter(abs(SOXmean_diff) >= abs(SOXtest_stat)) %>% summarise(n()/100)
SOX_p_CV <- as.numeric(SOX_moreextreme)

#SOX_p_CV

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Two sample hypothesis testing SOX, VEM+TRAM drug

SOXstarttimeCVT <- CVTpointfivetime %>% select(Sox10) #at 0.5 hours, #only 138 data points
SOXfulltimeCVT <- CVTfulltime %>% select(Sox10) #at 24 hours, has 5,000 data points

n1 <- 135
n2 <- 2000

set.seed(523)
N <- 1000; #how many times you run the code
SOXsimulated_values_CVT <- 1:N

for(i in 1:N){ #do 1000 'fake' observations
  SOXshuffled_CVT <- sample(c(SOXstarttimeCVT$Sox10, SOXfulltimeCVT$Sox10), size=n1+n2, replace=FALSE)
  SOXsimvalue_CVT <- mean(SOXshuffled_CVT[1:n1])-mean(SOXshuffled_CVT[(n1+1):(n1+n2)])
  SOXsimulated_values_CVT[i] <- as.numeric(SOXsimvalue_CVT)
}

simSOX_CVT <- tibble(SOXmean_diff_CVT = SOXsimulated_values_CVT)

#simSOX_CVT %>% ggplot(aes(x=SOXmean_diff_CVT)) + geom_histogram(binwidth=0.001, color="black", fill="gray")

SOXtest_stat_CVT <- 3.548716 - 3.539588

SOX_moreextremeCVT <- simSOX_CVT %>% filter(abs(SOXmean_diff_CVT) >= abs(SOXtest_stat_CVT)) %>% summarise(n()/100)
SOX_p_CVT <- as.numeric(SOX_moreextremeCVT); #SOX_p_CVT

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Two sample hypothesis testing NGFR, VEM drug

NGFRstarttimeCV <- CVpointfivetime %>% select(NGFR) #at 0.5 hours, #only 135 data points
NGFRfulltimeCV <- CVfulltime %>% select(NGFR) #at 24 hours, has 5,000 data points

n1 <- 135
n2 <- 4000

set.seed(523)
N <- 1000; #how many times you run the code
NGFRsimulated_values_CV <- 1:N

for(i in 1:N){ #do 1000 'fake' observations
  NGFRshuffled_CV <- sample(c(NGFRstarttimeCV$NGFR, NGFRfulltimeCV$NGFR), size=n1+n2, replace=FALSE)
  NGFRsimvalue_CV <- mean(NGFRshuffled_CV[1:n1])-mean(NGFRshuffled_CV[(n1+1):(n1+n2)])
  NGFRsimulated_values_CV[i] <- as.numeric(NGFRsimvalue_CV)
}

simNGFR <- tibble(NGFRmean_diff = NGFRsimulated_values_CV)

#simNGFR %>% ggplot(aes(x=NGFRmean_diff)) + geom_histogram(binwidth=0.001, color="black", fill="gray")

NGFRtest_stat <- 3.105181 - 3.131846

NGFR_moreextreme <- simNGFR %>% filter(abs(NGFRmean_diff) >= abs(NGFRtest_stat)) %>% summarise(n()/100)
NGFR_p_CV <- as.numeric(NGFR_moreextreme); #NGFR_p_CV

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Two sample hypothesis testing NGFR, VEM+TRAM drug

NGFRstarttimeCVT <- CVTpointfivetime %>% select(NGFR) #at 0.5 hours, #only 135 data points
NGFRfulltimeCVT <- CVTfulltime %>% select(NGFR) #at 24 hours, has 5,000 data points

n1 <- 195
n2 <- 2194

set.seed(523)
N <- 1000; #how many times you run the code
NGFRsimulated_values_CVT <- 1:N

for(i in 1:N){ #do 1000 'fake' observations
  NGFRshuffled_CVT <- sample(c(NGFRstarttimeCVT$NGFR, NGFRfulltimeCVT$NGFR), size=n1+n2, replace=FALSE)
  NGFRsimvalue_CVT <- mean(NGFRshuffled_CVT[1:n1])-mean(NGFRshuffled_CVT[(n1+1):(n1+n2)])
  NGFRsimulated_values_CVT[i] <- as.numeric(NGFRsimvalue_CVT)
}

simNGFR_CVT <- tibble(NGFRmean_diff_CVT = NGFRsimulated_values_CVT)

#simNGFR_CVT %>% ggplot(aes(x=NGFRmean_diff_CVT)) + geom_histogram(binwidth=0.001, color="black", fill="gray")

NGFRtest_stat_CVT <- 3.111332 - 3.145561

NGFR_moreextreme_CVT <- simNGFR_CVT %>% filter(abs(NGFRmean_diff_CVT) >= abs(NGFRtest_stat_CVT)) %>% summarise(n()/100)
NGFR_p_CVT <- as.numeric(NGFR_moreextreme_CVT); #NGFR_p_CVT


```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Two sample hypothesis testing MitFg, VEM drug

MITFGstarttimeCV <- CVpointfivetime %>% select(MiTFg) #at 0.5 hours, #only 135 data points
MITFGfulltimeCV <- CVfulltime %>% select(MiTFg) #at 24 hours, has 5,000 data points

n1 <- 135
n2 <- 4000

set.seed(523)
N <- 1000; #how many times you run the code
MITFGsimulated_values_CV <- 1:N

for(i in 1:N){ #do 1000 'fake' observations
  MITFGshuffled_CV <- sample(c(MITFGstarttimeCV$MiTFg, MITFGfulltimeCV$MiTFg), size=n1+n2, replace=FALSE)
  MITFGsimvalue_CV <- mean(MITFGshuffled_CV[1:n1])-mean(MITFGshuffled_CV[(n1+1):(n1+n2)])
  MITFGsimulated_values_CV[i] <- as.numeric(MITFGsimvalue_CV)
}

simMiTFg <- tibble(MITFGmean_diff = MITFGsimulated_values_CV)

#simMiTFg %>% ggplot(aes(x=MITFGmean_diff)) + geom_histogram(binwidth=0.001, color="black", fill="gray")

MiTFgtest_stat <- 3.105181 - 3.131846

MiTFg_moreextreme <- simMiTFg %>% filter(abs(MITFGmean_diff) >= abs(MiTFgtest_stat)) %>% summarise(n()/100)
MiTFg_p_CV <- as.numeric(MiTFg_moreextreme); #MiTFg_p_CV

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#Two sample hypothesis testing MitFg, VEM+TRAM drug

MITFGstarttimeCVT <- CVTpointfivetime %>% select(MiTFg) #at 0.5 hours, #only 135 data points
MITFGfulltimeCVT <- CVTfulltime %>% select(MiTFg) #at 24 hours, has 5,000 data points

n1 <- 135
n2 <- 2000

set.seed(523)
N <- 1000; #how many times you run the code
MITFGsimulated_values_CVT <- 1:N

for(i in 1:N){ #do 1000 'fake' observations
  MITFGshuffled_CVT <- sample(c(MITFGstarttimeCVT$MiTFg, MITFGfulltimeCVT$MiTFg), size=n1+n2, replace=FALSE)
  MITFGsimvalue_CVT <- mean(MITFGshuffled_CVT[1:n1])-mean(MITFGshuffled_CVT[(n1+1):(n1+n2)])
  MITFGsimulated_values_CVT[i] <- as.numeric(MITFGsimvalue_CVT)
}

simMiTFgCVT <- tibble(MITFGmean_diff_CVT = MITFGsimulated_values_CVT)

#simMiTFgCVT %>% ggplot(aes(x=MITFGmean_diff_CVT)) + geom_histogram(binwidth=0.001, color="black", fill="gray")

MiTFgtest_stat_CVT <- 3.835648 - 3.782575

MiTFg_moreextreme_CVT <- simMiTFgCVT %>% filter(abs(MITFGmean_diff_CVT) >= abs(MiTFgtest_stat_CVT)) %>% summarise(n()/100)
MiTFg_p_CVT <- as.numeric(MiTFg_moreextreme_CVT); #MiTFg_p_CVT

```

## Resulting P-Values for Cancerous Cells

### From the two-sample hypothesis tests, our main findigns were;

- AXL_VEM drug → P-value was so small, it showed up as 0 on the code.\
- AXL_VEM+TRAM drug → P-value was so small, it showed up as 0 on the code.\
- NGFR_VEM drug → P-value was so small, it showed up as 0 on the code.\
- NGFR_VEM+TRAM drug → P-value was so small, it showed up as 0 on the code.\
- MiTFg_VEM drug → P-value was so small, it showed up as 0 on the code.\
- MiTFg_VEM+TRAM drug → P-value was so small, it showed up as 0 on the code.\
- Sox10 VEM drug → P-value was 0.74. Fail to reject the null hypothesis.\
- Sox10 VEM+TRAM drug → P-value was 0.36. Fail to reject the null hypothesis.\

## Correlation Estimation 

**What is the relationship between different proteins at a fixed timepoint and environmental condition?**\
- We wanted to investigate this as it is an essential component to understanding how phenotypes behave and what environmental conditions, such as other phenotypes, can affect their own phenotype level.\
- If we find trends, we can use this to justify further scientific exploration into how phenotypes themselves can be used to reduce melanocytic cells.\

---

## Correlation Estimation

```{r, echo=FALSE, message=FALSE, warning=FALSE}

install.packages("cowplot")
library("cowplot")


```


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3.1, fig.width=6}
dataset_cancer<- read_csv("STA130Project.csv")
dataset_cancer%>% na.omit() %>% arrange() -> dataset_cancer


median_value_Sox10<-as.numeric(dataset_cancer%>%summarise(median(Sox10)))
median_value_AXL<-as.numeric(dataset_cancer%>%summarise(median(AXL)))
median_value_NGFR<-as.numeric(dataset_cancer%>%summarise(median(NGFR)))
median_value_MiTFg<-as.numeric(dataset_cancer%>%summarise(median(MiTFg)))

dataset_cancer %>% mutate(NumValue_Sox10=case_when(median_value_Sox10<Sox10~1,median_value_Sox10>=Sox10~0)) %>% mutate(NumValue_AXL=case_when(median_value_AXL<Sox10~1,median_value_AXL>=AXL~0))%>% mutate(NumValue_NGFR=case_when(median_value_NGFR<NGFR~1,median_value_NGFR>=NGFR~0))%>% mutate(NumValue_MiTFg=case_when(median_value_MiTFg<Sox10~1,median_value_MiTFg>=MiTFg~0))%>% mutate(Sum_of_Proteins=NumValue_MiTFg+NumValue_NGFR+NumValue_AXL+NumValue_Sox10)%>% mutate(Cancer_Binary_Evaluator=case_when(Sum_of_Proteins==1~"Undifferentiated",Sum_of_Proteins==3~"Neural crest-like or Transitory",Sum_of_Proteins==2~"Melanocytic")) %>% mutate(Cancer=case_when(Cancer_Binary_Evaluator=="Undifferentiated"~"Negative",Cancer_Binary_Evaluator=="Melanocytic"~"Positive")) %>% na.omit() ->dataset_cancer






datapoint <- dataset_cancer %>% filter(Drugs=='Vem') %>% filter(Timepoint=='0.5 h')





cor(datapoint$Sox10, datapoint$AXL)^2
cor(datapoint$MiTFg, datapoint$NGFR)^2

cor(datapoint$AXL, datapoint$NGFR)^2
cor(datapoint$AXL, datapoint$MiTFg)^2


cor(datapoint$Sox10, datapoint$NGFR)^2
cor(datapoint$MiTFg, datapoint$Sox10)^2



a<-datapoint %>%
ggplot(aes(x=AXL, y=Sox10)) +
geom_point(aes(color=Cancer)) + theme_light() +
labs(title="AXL vs Sox10",
x="AXL",
y="Sox10")  + theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  ) + theme(plot.background = element_rect(fill = "purple"))


b<-datapoint %>%
ggplot(aes(x=NGFR, y=Sox10)) +
geom_point(aes(color=Cancer)) + theme_light() +
labs(title="NGFR vs Sox10",
x="AXL",
y="Sox10")  + theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  ) + theme(plot.background = element_rect(fill = "purple"))


c<-datapoint %>%
ggplot(aes(x=MiTFg, y=Sox10)) +
geom_point(aes(color=Cancer)) + theme_light() +
labs(title="MitFg vs Sox10",
x="AXL",
y="Sox10")  + theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  ) + theme(plot.background = element_rect(fill = "purple"))




d<-datapoint %>%
ggplot(aes(x=NGFR, y=MiTFg)) +
geom_point(aes(color=Cancer)) + theme_light() +
labs(title="NGFR vs MiTFg",
x="NGFR",
y="MiTFg")  + theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  ) + theme(plot.background = element_rect(fill = "purple"))


e<-datapoint %>%
ggplot(aes(x=AXL, y=MiTFg)) +
geom_point(aes(color=Cancer)) + theme_light() +
labs(title="AXL vs MiTFg",
x="AXL",
y="Sox10")  + theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  ) + theme(plot.background = element_rect(fill = "purple"))



f<-datapoint %>%
ggplot(aes(x=AXL, y=NGFR)) +
geom_point(aes(color=Cancer)) + theme_light() +
labs(title="AXL vs NGFR",
x="AXL",
y="Sox10")  + theme(
  panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")
  ) + theme(plot.background = element_rect(fill = "purple"))


```

---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3.1, fig.width=6}

plot_grid(a,b,c,d,e,f)

```

---

## BOOTSTRAPPING CORRELATION ESTIMATES FOR CONFIDENCE INTERVALS

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=1.5, fig.width=1.6}


#AXL vs Sox10

N = 100
dataset_cancer%>% na.omit()-> dataset_cancer
#print(correlation_value_storage[i] <- cor(dataset_cancer$AXL,dataset_cancer$Sox10)^2)


correlation_value_storage <- rep(NA,nrow(dataset_cancer))
#  data_shuffled<- df %>% sample_n(5)
set.seed(223)

for(i in seq_along(1:N)) {

  data_shuffled <-  sample_n(dataset_cancer,size=10000,replace=TRUE)


  correlation_value_storage[i] <- cor(data_shuffled$AXL,data_shuffled$Sox10)^2
  #print(correlation_value_storage[i])
}
bootstraped_data<-correlation_value_storage%>% na.omit()
table1 <- tibble(values=bootstraped_data)
a<-ggplot(table1,aes(x=values))+geom_histogram(bins=16,color="blue",fill="green")+labs(title="AXL vs NGFR",
                                                                                       x="AXL",
                                                                                       y="Sox10")






#MiTFg vs  Sox10
N = 100
#print(correlation_value_storage[i] <- cor(dataset_cancer$AXL,dataset_cancer$Sox10)^2)


correlation_value_storage <- rep(NA,nrow(dataset_cancer))
#  data_shuffled<- df %>% sample_n(5)
set.seed(223)

for(i in seq_along(1:N)) {

  data_shuffled <-  sample_n(dataset_cancer,size=10000,replace=TRUE)


  correlation_value_storage[i] <- cor(data_shuffled$MiTFg,data_shuffled$Sox10)^2
  #print(correlation_value_storage[i])
}
bootstraped_data<-correlation_value_storage%>% na.omit()
table2 <- tibble(values=bootstraped_data)
b<-ggplot(table2,aes(x=values))+geom_histogram(bins=16,color="blue",fill="green")+labs(title="MiTFg vs NGFR",
                                                                                       x="MiTFg",
                                                                                       y="Sox10")








#NGFR vs Sox10
N = 100
#print(correlation_value_storage[i] <- cor(dataset_cancer$AXL,dataset_cancer$Sox10)^2)


correlation_value_storage <- rep(NA,nrow(dataset_cancer))
#  data_shuffled<- df %>% sample_n(5)
set.seed(223)

for(i in seq_along(1:N)) {

  data_shuffled <-  sample_n(dataset_cancer,size=10000,replace=TRUE)


  correlation_value_storage[i] <- cor(data_shuffled$NGFR,data_shuffled$Sox10)^2
  #print(correlation_value_storage[i])
}
bootstraped_data<-correlation_value_storage%>% na.omit()
table3 <- tibble(values=bootstraped_data)
 c<-ggplot(table3,aes(x=values))+geom_histogram(bins=16,color="blue",fill="green")+labs(title="NGFR vs Sox10",
                                                                                       x="NGFR",
                                                                                       y="Sox10")




#NGFR vs MiTFg

N = 100
#print(correlation_value_storage[i] <- cor(dataset_cancer$AXL,dataset_cancer$Sox10)^2)


correlation_value_storage <- rep(NA,nrow(dataset_cancer))
#  data_shuffled<- df %>% sample_n(5)
set.seed(223)

for(i in seq_along(1:N)) {

  data_shuffled <-  sample_n(dataset_cancer,size=10000,replace=TRUE)


  correlation_value_storage[i] <- cor(data_shuffled$NGFR,data_shuffled$MiTFg)^2
  #print(correlation_value_storage[i])
}
bootstraped_data<-correlation_value_storage%>% na.omit()
table4 <- tibble(values=bootstraped_data)
d<-ggplot(table4,aes(x=values))+geom_histogram(bins=16,color="blue",fill="green")+labs(title="NGFR vs MiTFg",
                                                                                      x="NGFR",
                                                                                      y="MiTFg")




#AXL vs MiTFg
N = 100
#print(correlation_value_storage[i] <- cor(dataset_cancer$AXL,dataset_cancer$Sox10)^2)


correlation_value_storage <- rep(NA,nrow(dataset_cancer))
#  data_shuffled<- df %>% sample_n(5)
set.seed(223)

for(i in seq_along(1:N)) {

  data_shuffled <-  sample_n(dataset_cancer,size=10000,replace=TRUE)


  correlation_value_storage[i] <- cor(data_shuffled$AXL,data_shuffled$MiTFg)^2
  #print(correlation_value_storage[i])
}
bootstraped_data<-correlation_value_storage%>% na.omit()
table5<- tibble(values=bootstraped_data)
e<-ggplot(table5,aes(x=values))+geom_histogram(bins=16,color="blue",fill="green")+labs(title="AXL vs MiTFg",
                                                                                      x="AXL",
                                                                                      y="MiTFg")






#AXL  vs y=NGFR
N = 100
#print(correlation_value_storage[i] <- cor(dataset_cancer$AXL,dataset_cancer$Sox10)^2)


correlation_value_storage <- rep(NA,nrow(dataset_cancer))
#  data_shuffled<- df %>% sample_n(5)
set.seed(223)

for(i in seq_along(1:N)) {

  data_shuffled <-  sample_n(dataset_cancer,size=10000,replace=TRUE)


  correlation_value_storage[i] <- cor(data_shuffled$AXL,data_shuffled$NGFR)^2
  #print(correlation_value_storage[i])
}
bootstraped_data<-correlation_value_storage%>% na.omit()
table6 <- tibble(values=bootstraped_data)
f<-ggplot(table6,aes(x=values))+geom_histogram(bins=16,color="blue",fill="green")+labs(title="AXL vs NGFR",
                                                                                      x="AXL",
                                                                                      y="NGFR")



a+theme(axis.text.x=element_text(size=5))
b+theme(axis.text.x=element_text(size=5))
c+theme(axis.text.x=element_text(size=5))
d+theme(axis.text.x=element_text(size=5))
e+theme(axis.text.x=element_text(size=5))
f+theme(axis.text.x=element_text(size=5))


plot_grid(a,b,c,d,e,f)


```

## Main Results

### Our main results were...

From the classification trees, it can be concluded that;\
\vspace{0.3cm}
-  If we know the level of the TFs Phospho_Fra1, Fra2, Phospho_Erk1, Phospho_ATF2, ATF6, Ki_67, c_Fos, ATF4, Phospho_S6, and ATF3, then we can tell if each protein level is HIGH or LOW. From that, we combine our results and figure out if the cell is cancerous or not.\
\vspace{0.5cm}
From the two-sample hypothesis tests, it can be concluded that:\
\vspace{0.3cm}
- There is a significant change overtime for the AXL, NGFR and MiTFg proteins in the VEM and VEM+TRAM drug.\
- In the VEM drug, AXL decreases overtime, NGFR increases overtime and MiTFg increases overtime.\
- In the VEM+TRAM drug, AXL decreases overtime, NGFR increases overtime and MiTFg decreases overtime.\
- We fail to reject the null hypothesis for Sox10.\
\vspace{0.5cm}
From the correlation estimation and bootstrapping, it can be concluded that;\
- There exists a weak correlation between different proteins which may force us to reassess the importance of the relative quantities of proteins.\


## Conclusion

### We have concluded that...

- In the presence of the VEM+TRAM drug, MiTFg levels reduce. This is a significant finding since we know cells are cancerous when they have high MiTFg, high NGFR, low Sox10 and low AXL.\
- Although Sox10 is relatively unchanged overtime, AXL further decreases and NGFR further increases, the significant change of one protein away from the regular cancerous protein configuration is promising and provides a scientific basis to further research how proteins respond to the Vem and Tram drug.\
- Furthermore, given the weak correlation between key phenotypes, the findings from this investigation provide a basis for further scientific investigation into how phenotypes and various drugs may reduce the development of melanocytic cells.

## Considerations

### Some considerations include;

- Each experiment was made under the condition that everything else remained constant except the variables under consideration.\
- Sox10 has a noticable change in protein levels at the 2h timepoint which, due to the limitations of linear regression, could not be aptly included in this investigation.\

### Acknowledgements

All code and functions used during this investigation come from the following R libraries; tidyverse, rpart, partykit and cowplot.
